### 2023-05-06 manuscript/paper.Rmd

---
title: "test"
author: 
  - "Angelos Vasilopoulos"
  - "Gregory J. Matthews"
abstract: |
  | Abstract \vspace{2mm}
  | *Keywords*: Cross validation  
bibliography: references.bib
fontsize: 12pt
link-citations: true
linkcolor: cyan
urlcolor: cyan
output:
  pdf_document:
    df_print: kable
    number_sections: true
    keep_tex: true
header-includes:
 \usepackage{setspace}
 \setstretch{1.15}
 \usepackage{float}
 \floatplacement{figure}{t}
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)

gjm <- function(x, color = "red") {
  sprintf("\\textcolor{%s}{%s}", color, x)}
```

\newpage

# Introduction {#sec:intro}
We wanted to look at how the choice of k in k-fold cross validation affected the estimation of the prediction error.

Literature review.

# Simulation Study
This is how you cite a paper: @R2022language

Simulations were performed using the R programming language (@R2022language).

Describe simulation studty with LASSO.

# Results
We find blah blah

#Conclusions 
Early work found this.  More recent work found this.  Our work confirms that.  We suggest........

Future work?  What else would we like to do?

# Acknowledgements {-}

Everyone who took STAT338 in Fall 2022.

# Supplementary Material {-}

All code for reproducing the analyses in this paper is publicly available at https://github.com/gjm112/optimal_k

# References